Intergroup contact fosters\\more inclusive social identities

### 2018-11-09 manuscript/manuscript.tex
prepare for submission

% Format
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0}
\usepackage{lineno}
%\usepackage{authblk}

@@

%\linenumbers

@@

We tested these predictions in a triple crossed-categorisation task (adapted from \citeNP{dommelen_construing_2015}). In this task, participants viewed identity cards, each representing a person with whom participants shared none, one, two, or all of three group memberships. Participants reported whether they considered each person as ``us'' or ``not us'', showing whom they included in, or excluded from, their ingroup. This task thus measured how participants combined multiple group memberships to construct their social identities. We tested to what extent participants' responses were associated with past contact experiences, ideological preferences, and various outcome measures.

@@

\begin{figure}
\centering
\includegraphics[scale=1]{../figures/figure-2}
\caption{Examples of targets used in the triple crossed-categorization task. Based on ratings in a pilot study ($N = 26$), we selected the four most prototypical targets (out of fifty initial targets) for each of six plausible combinations of caste, religion, and nationality (for details, see Appendix~A). Each target showed a person's caste (GM = General Merit, OBC = Other Backward Class, SC/ST = Scheduled Caste/Scheduled Tribe), religion (Hindu, Muslim), and nationality (Indian, Nepali, Sri Lankan, Bangladeshi). Each target also showed the person's first and last name, age (21--26 years), and a silhouette corresponding to the person's gender (adapted from \protect\citeNP{ma_chicago_2015}). Each target's age and silhouette, as well as the order in which the targets were presented, varied across sessions.}
\label{fig:f2}
\end{figure}

@@

We compared \emph{how likely} participants were to categorize \emph{which} target as ``us'' versus ``not us''---and how that probability varied across targets' and participants' group memberships.\footnote{In Appendix~B, we report analyses using \citeauthor{dommelen_construing_2015}'s \citeyear{dommelen_construing_2015} operationalisation who analysed which and how many targets participants included as distinct questions. }

@@

\begin{table}
\caption{
Comparison of models estimating the probability of participants categorising targets as ``us'' versus ``not us''. $\textit{ELPD}$ is the expected log predictive density, with higher numbers indicating that a model is expected to make more accurate out-of-sample predictions \protect\fullcite{vehtari_practical_2017}. $\Delta\textit{ELPD}$ is the difference in $\textit{ELPD}$ between the current and previous model, with positive values indicating that the current model is expected to make more accurate out-of-sample predictions. We selected a more complex model over a simpler model when $\frac{\Delta\textit{ELPD}}{\textit{SE}} \geq 2$. $w$ are stacking weights based on the models' expected log predictive densities \protect\cite{yao_using_2018}.
}
\centering
\figureversion{lining, tabular}
\small
\begin{tabularx}{\linewidth}{r@{~}rXrrrrrr} \toprule
\# &  & Description & $\textit{ELPD}$ & $\textit{SE}$ & $\Delta\textit{ELPD}$ & $\textit{SE}$ & $\frac{\Delta\textit{ELPD}}{\textit{SE}}$ & $w$ \\ \midrule 
0 &      & Intercept (Participant) & -4703.5 & 23.7 &     - &    - &    - & .00 \\ 
1 & vs 0 & Intercept (Category)    & -3853.5 & 41.6 & 850.1 & 37.4 & 22.7 & .00 \\
2 & vs 1 & Group differences (SC/ST)       & -3791.3 & 42.0 &  62.1 & 11.1 &  5.6 & .00 \\
3 & vs 2 & Group differences (OBC)         & -3792.1 & 42.2 &  -0.8 &  3.6 & -0.2 & .07 \\ \midrule
4 & vs 2 & Intergroup contact (4) & -3743.4 & 42.8 &  47.9 &  8.8 &  5.4 & .00 \\
5 & vs 4 & Intergroup contact (2) & -3736.7 & 42.6 &   6.7 &  1.8 &  3.8 & .92 \\
6 & vs 5 & Intergroup contact (2) & -3747.8 & 42.7 & -11.1 &  3.2 & -3.5 & .00 \\
7 & vs 2 & Social dominance orientation & -3791.0 & 42.1 &   0.4 &  3.3 &  0.1 & .00 \\
\bottomrule
\end{tabularx}
\label{tab:t2}
\end{table}

@@

As expected, participants' caste membership shaped how they categorized Indians of different castes and religions. Most GM and OBC participants included \emph{Hindu, GM} and \emph{Hindu, OBC} targets in their ingroup, \pest{M2}{.94}{.92}{.96} and \pest{M2}{.93}{.91}{.95}. Fewer GM and OBC participants categorized \emph{Hindu, SC/ST} targets as ``us'', \pest{M2}{.84}{.80}{.87}. They were least likely to categorize \emph{Muslim, OBC} targets as ``us'', \pest{M2}{.75}{.70}{.80}. Dalit/Adivasi (SC/ST) participants' responses differed from GM and OBC participants'. Almost all SC/ST participants included \emph{Hindu, SC/ST} targets in their ingroup, \pest{M2}{.97}{.95}{.99}. Fewer SC/ST participants included \emph{Hindu, GM} and \emph{Hindu, OBC} in their ingroup, \pest{M2}{.88}{.83}{.92} and  and \pest{M2}{.80}{.73}{.86}. SC/ST participants were less likely than others to include \emph{Muslim, OBC} targets as ``us'', \pest{M2}{.47}{.38}{.57}. These findings showed that participants from advantaged backgrounds tended to exclude targets from disadvantaged backgrounds (and vice versa), while all participants tended to exclude targets from the Muslim minority.

@@

\begin{figure}
\centering
\includegraphics[scale=1]{../figures/figure-4}
\caption{Estimated probability of participants categorising a target as ``us'' versus ``not us'' as a function of the targets' group memberships (horizontal), the participants' group memberships (colour), and the reported amount of negative contact and outgroup friendship with the relevant groups (in Model~5, Table~\ref{tab:t2}). GM = General Merit, OBC = Other Backward Class, SC/ST = Scheduled Caste/Scheduled Tribe.}
\label{fig:f4}
\end{figure}

@@

We analysed how participants' categorizations related to their warmth and social distance ratings for each target in the categorization task, to intergroup threat, and to their perceived (dis-)advantage and policy support.

@@

Our findings show that the triple crossed-categorization task \cite{dommelen_construing_2015} is an intuitive and informative method for studying social identification across multiple categories. We adapted the task to answer new questions about social identification. First, we recruited respondents from multiple groups, allowing us to study both individual and group differences (see also \citeNP{brankovic_social_2015}). Second, we estimated responses as varying across targets and participants using multilevel models. This allowed more fine-grained analyses than \citeauthor{dommelen_construing_2015}’s quantitative and qualitative summaries, and may explain why we found more consistent effects of intergroup contact. Overall, these changes open the triple crossed-categorization task to a broader range of research questions.

>>>>

% Format
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0}
\usepackage[pagewise]{lineno}
%\usepackage{authblk}

@@

\linenumbers

@@

We tested these predictions in a triple crossed-categorization task (adapted from \citeNP{dommelen_construing_2015}). In this task, participants viewed identity cards, each representing a person with whom participants shared none, one, two, or all of three group memberships. Participants reported whether they considered each person as ``us'' or ``not us'', showing whom they included in, or excluded from, their ingroup. This task thus measured how participants combined multiple group memberships to construct their social identities. We tested to what extent participants' responses were associated with past contact experiences, ideological preferences, and various outcome measures.

@@

\begin{figure}
\centering
\includegraphics[scale=1]{../figures/figure-2}
\caption{Examples of targets used in the triple crossed-categorization task. Based on ratings in a pilot study ($N = 26$), we selected the four most prototypical targets (out of fifty initial targets) for each of six plausible combinations of caste, religion, and nationality (for details, see Appendix~A). Each target showed a person's caste reservation (GM = General Merit, OBC = Other Backward Class, SC/ST = Scheduled Caste/Scheduled Tribe), religion (Hindu, Muslim), and nationality (Indian, Nepali, Sri Lankan, Bangladeshi). Each target also showed the person's first and last name, age (21--26 years), and a silhouette corresponding to the person's gender (adapted from \protect\citeNP{ma_chicago_2015}). Each target's age and silhouette, as well as the order in which the targets were presented, varied across sessions.}
\label{fig:f2}
\end{figure}

@@

We compared \emph{how likely} participants were to categorize \emph{which} target as ``us'' versus ``not us''---and how that probability varied across targets' and participants' group memberships.\footnote{In Appendix~B, we report analyses using \citeauthor{dommelen_construing_2015}'s \citeyear{dommelen_construing_2015} operationalisation; they analysed \emph{which} and \emph{how many} targets participants included as distinct questions.}

@@

\begin{table}
\caption{
Comparison of models estimating the probability of participants categorizing targets as ``us'' versus ``not us''. $\textit{ELPD}$ is the expected log predictive density, with higher numbers indicating that a model is expected to make more accurate out-of-sample predictions \protect\fullcite{vehtari_practical_2017}. $\Delta\textit{ELPD}$ is the difference in $\textit{ELPD}$ between the current and previous model, with positive values indicating that the current model is expected to make more accurate out-of-sample predictions. We selected a more complex model over a simpler model when $\frac{\Delta\textit{ELPD}}{\textit{SE}} \geq 2$. $w$ are stacking weights based on the models' expected log predictive densities \protect\cite{yao_using_2018}.
}
\centering
\figureversion{lining, tabular}
\small
\begin{tabularx}{\linewidth}{r@{~}rXrrrrrr} \toprule
\# &  & Description & $\textit{ELPD}$ & $\textit{SE}$ & $\Delta\textit{ELPD}$ & $\textit{SE}$ & $\frac{\Delta\textit{ELPD}}{\textit{SE}}$ & $w$ \\ \midrule 
0 &      & Intercept (Participant) & -4703.5 & 23.7 &     - &    - &    - & .00 \\ 
1 & vs 0 & Intercept (Category)    & -3853.5 & 41.6 & 850.1 & 37.4 & 22.7 & .00 \\
2 & vs 1 & Group differences (SC/ST)       & -3791.3 & 42.0 &  62.1 & 11.1 &  5.6 & .00 \\
3 & vs 2 & Group differences (OBC)         & -3792.1 & 42.2 &  -0.8 &  3.6 & -0.2 & .07 \\ \midrule
4 & vs 2 & Intergroup contact (4) & -3743.4 & 42.8 &  47.9 &  8.8 &  5.4 & .00 \\
5 & vs 4 & Intergroup contact (2) & -3736.7 & 42.6 &   6.7 &  1.8 &  3.8 & .92 \\
6 & vs 5 & Intergroup contact (2) & -3747.8 & 42.7 & -11.1 &  3.2 & -3.5 & .00 \\
7 & vs 2 & Social dominance orientation & -3791.0 & 42.1 &   0.4 &  3.3 &  0.1 & .00 \\
\bottomrule
\end{tabularx}
\label{tab:t2}
\end{table}

@@

As expected, participants' caste membership shaped how they categorized Indians of different castes and religions. Most GM and OBC participants included \emph{Hindu, GM} and \emph{Hindu, OBC} targets in their ingroup, \pest{M2}{.94}{.92}{.96} and \pest{M2}{.93}{.91}{.95}. Fewer GM and OBC participants categorized \emph{Hindu, SC/ST} targets as ``us'', \pest{M2}{.84}{.80}{.87}. They were least likely to categorize \emph{Muslim, OBC} targets as ``us'', \pest{M2}{.75}{.70}{.80}. Dalit/Adivasi (SC/ST) participants' responses differed from GM and OBC participants'. Almost all SC/ST participants included \emph{Hindu, SC/ST} targets in their ingroup, \pest{M2}{.97}{.95}{.99}. Fewer SC/ST participants included \emph{Hindu, GM} and \emph{Hindu, OBC} targets in their ingroup, \pest{M2}{.88}{.83}{.92} and  and \pest{M2}{.80}{.73}{.86}. SC/ST participants were less likely than others to include \emph{Muslim, OBC} targets as ``us'', \pest{M2}{.47}{.38}{.57}. These findings showed that participants from advantaged backgrounds tended to exclude targets from disadvantaged backgrounds (and vice versa), while all participants tended to exclude targets from the Muslim minority.

@@

\begin{figure}
\centering
\includegraphics[scale=1]{../figures/figure-4}
\caption{Estimated probability of participants categorizing a target as ``us'' versus ``not us'' as a function of the targets' group memberships (horizontal), the participants' group memberships (colour), and the reported amount of negative contact and outgroup friendship with the relevant groups (in Model~5, Table~\ref{tab:t2}). GM = General Merit, OBC = Other Backward Class, SC/ST = Scheduled Caste/Scheduled Tribe.}
\label{fig:f4}
\end{figure}

@@

We analysed how participants' categorizations related to their feeling thermometerhttps://www.dur.ac.uk/jobs/recruitment/vacancies/psyc18-10/Over and social distance ratings for each target in the categorization task, to intergroup threat, and to their perceived (dis-)advantage and policy support.

@@

Our findings show that the triple crossed-categorization task \cite{dommelen_construing_2015} is an intuitive and informative method for studying social identification across multiple categories. We adapted the task to answer new questions about social identification. First, we recruited respondents from multiple groups, allowing us to study both individual and group differences (see also \citeNP{brankovic_social_2015}). Second, we estimated responses as varying across targets and participants using multilevel models. This allowed more fine-grained analyses than \citeauthor{dommelen_construing_2015}’s quantitative and qualitative summaries, and may explain why we found more consistent effects of intergroup contact. Together, these changes open the triple crossed-categorization task to a broader range of research questions.
